The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-05-08 writeup.Rmd
edits to introduction

The most prominently studied phenomenon in experimental pragmatics is undoubtedly *scalar implicature*. Scalar implicatures arise in virtue of a speaker producing the weaker of two ordered scalemates [@grice1975; @Horn1972; @Hirschberg1985; @Geurts2010]. Examples are provided in (1-2).

@@

A standard measure that has stood proxy for implicature rate across many studies is the proportion of 'pragmatic' judgments in truth-value judgment paradigms [@Bott2004;@Noveck2001;@Noveck2003;@Chemla2011;@Geurts2009;@DegenTanenhaus2015;@DeNeys2007;@Degen2014]. In these kinds of tasks, participants are provided a set of facts, either presented visually or via their own knowledge of the world. They are then asked to judge whether a sentence intended to describe those facts is true or false (or alternatively, whether it is right or wrong, or they are asked whether they agree or disagree with the sentence). The crucial condition for assessing implicature rates in these kinds of studies typically consists of a case where the facts are such that the stronger alternative is true and the target utterance is thus also true but underinformative. For instance, @Bott2004 asked participants to judge sentences like 'Some elephants are mammals', when world knowledge dictates that all elephants are mammals. Similarly, @DegenTanenhaus2015 asked participants to judge sentences like 'You got some of the gumballs' in situations where the visual evidence indicated that the participant received all the gumballs from a gumball machine. In these kinds of scenarios, the story goes, if a participant responds 'FALSE', that indicates that they computed a scalar implicature, eg to the effect of 'Not all elephants are mammals' or 'You didn't get all of the gumballs', which is (globally or contextually) false. If instead a participant responds 'TRUE', that is taken to indicate that they interpreted the utterance literally as `Some, and possibly all, elephants are mammals' or 'You got some, and possibly all, of the gumballs'.

@@

We begin by reporting an experiment that addresses the following question: do the number of response options provided in a truth-value judgment task and the way that responses are grouped into pragmatic ('SI') and literal ('no SI') change inferences about scalar implicature rates? Note that this way of asking the question presupposes two things: first, that whatever participants are doing in a truth-value judgment task, the behavioral measure can be interpreted as providing a measure of *interpretation*. And second, that listeners either do or do not compute an implicature on any given occasion. In the General Discussion we will discuss both of these issues. First, following @Degen2014, we will offer some remarks on why truth-value judgment tasks are better thought of as measuring participants' estimates of speakers' *production* probabilities. This will suggest a completely different class of linking functions. And second, we discuss an alternative conception of scalar implicature as a probabilistic phenomeonen, a view that has recently rose to prominence in the subfield of probabilistic pragmatics [@Franke2016;@Goodman2016]. This alternative conception of scalar implicature, we argue, affords developing and testing quantitative linking functions in a rigorous and motivated way.

@@

Let's examine an example. Assume three response options (wrong, neither, right). Assume further that a third of participants each gave each of the three responses, i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are for instance interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup. We manipulated the number of response options between participants and analyzed the results under different linking assumptions.

>>>>

The most prominently studied phenomenon in experimental pragmatics is undoubtedly *scalar implicature*. Scalar implicatures arise in virtue of **bw: as a result of?** a speaker producing the weaker of two ordered scalemates [@grice1975; @Horn1972; @Hirschberg1985; @Geurts2010]. Examples are provided in (1-2).

@@

A standard measure that has stood proxy **bw: stood as a proxy** for implicature rate across many studies is the proportion of 'pragmatic' judgments in truth-value judgment paradigms [@Bott2004;@Noveck2001;@Noveck2003;@Chemla2011;@Geurts2009;@DegenTanenhaus2015;@DeNeys2007;@Degen2014]. In these kinds of tasks, participants are provided a set of facts, either presented visually or via their own knowledge of the world. They are then asked to judge whether a sentence intended to describe those facts is true or false (or alternatively, whether it is right or wrong, or they are asked whether they agree or disagree with the sentence). The crucial condition for assessing implicature rates in these kinds of studies typically consists of a case where the facts are such that the stronger alternative is true and the target utterance is thus also true but underinformative. For instance, @Bott2004 asked participants to judge sentences like 'Some elephants are mammals', when world knowledge dictates that all elephants are mammals. Similarly, @DegenTanenhaus2015 asked participants to judge sentences like 'You got some of the gumballs' in situations where the visual evidence indicated that the participant received all the gumballs from a gumball machine. In these kinds of scenarios, the story goes, if a participant responds 'FALSE', that indicates that they computed a scalar implicature, eg to the effect of 'Not all elephants are mammals' or 'You didn't get all of the gumballs', which is (globally or contextually) false. If instead a participant responds 'TRUE', that is taken to indicate that they interpreted the utterance literally as `Some, and possibly all, elephants are mammals' or 'You got some, and possibly all, of the gumballs'.

@@

We begin by reporting an experiment that addresses the following question: do the number of response options provided in a truth-value judgment task and the way that responses are grouped into pragmatic ('SI') and literal ('no SI') change inferences about scalar implicature rates? Note that this way of asking the question presupposes two things: first, that whatever participants are doing in a truth-value judgment task, the behavioral measure can be interpreted as providing a measure of *interpretation*. And second, **bw: change to semicolon, i.e. "... interpretation; and second,**  that listeners either do or do not compute an implicature on any given occasion. In the General Discussion we will discuss both of these issues. First, following @Degen2014, we will offer some remarks on why truth-value judgment tasks are better thought of as measuring participants' estimates of speakers' *production* probabilities. This will suggest a completely different class of linking functions. And second, **bw: Next, (to avoid redundancy -- see above)** we discuss an alternative conception of scalar implicature as a probabilistic phenomeonen, a view that has recently rose to prominence in the subfield of probabilistic pragmatics [@Franke2016;@Goodman2016]. This alternative conception of scalar implicature, we argue, affords developing and testing quantitative linking functions in a rigorous and motivated way.

@@

Let's examine an example. Assume three response options (wrong, neither, right). Assume further that a third of participants each gave each of the three responses, **bw: Assume further that each of the three responses was selected by a third of participants,** i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are for instance **bw: if we are, for instance, **interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup. We manipulated the number of response options between participants and analyzed the results under different linking assumptions.
